---
title: "KNN"
author: "Su Goh"
date: "02/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, load the FNN package to use KNN methods. I'm going to use the Boston dataset, and use lower status of the population (percent) (`lstat`) to predict per capita crime rate by town (`crim`).
```{r load}
library(FNN)
library(MASS)
x =  Boston$lstat
y = Boston$crim
```

Plot the observations:
```{r plot}
plot(x, y, xlab = "Lower status of population (%)", ylab = "Per capita crime rate")
```

To use knn, we need a training set and a test set. The test set is the predictor values, x, at which we would like to make predictions. 

Here, the test set we create is a sequence of numbers between the min and max of the x predictors, with increments by 0.01
```{r test}
lstat_grid = data.frame(lstat = seq(min(Boston$lstat), max(Boston$lstat), by=0.01))
# or equivalently
# lstat_grid = data.frame(lstat = seq(range(Boston$lstat)[1], range(Boston$lstat)[2], by=0.01))
```

Now, we can make knn predictions, using the `knn.reg()` function from the FNN package. For the sake of the exercise, we'll try different values of k.
```{r pred}
pred_k1 = knn.reg(train=x, test=lstat_grid, y=y, k=1)
pred_k5 = knn.reg(train=x, test=lstat_grid, y=y, k=5)
pred_k10 = knn.reg(train=x, test=lstat_grid, y=y, k=10)
pred_k50 = knn.reg(train=x, test=lstat_grid, y=y, k=50)
pred_k100 = knn.reg(train=x, test=lstat_grid, y=y, k=100)
n = dim(Boston)[1]
pred_kn = knn.reg(train=x, test=lstat_grid, y=y, k=n)
```

View these predictions:
```{r pred plot}
par(mfrow = c(3,2))

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=1")
lines(x = lstat_grid$lstat, y = pred_k1$pred, col='orange')

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=5")
lines(x = lstat_grid$lstat, y = pred_k5$pred, col='orange')

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=10")
lines(x = lstat_grid$lstat, y = pred_k10$pred, col='orange')

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=50")
lines(x = lstat_grid$lstat, y = pred_k50$pred, col='orange')

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=100")
lines(x = lstat_grid$lstat, y = pred_k100$pred, col='orange')

plot(y ~ x, xlab = "Lower status of population (%)", ylab = "Per capita crime rate", main="k=n")
lines(x = lstat_grid$lstat, y = pred_kn$pred, col='orange')
```

