---
title: "SLR Simulation"
author: "Su Goh"
date: "09/10/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

addTrans <- function(color,trans) {
  # This function adds transparancy to a color.
  # Define transparancy with an integer between 0 and 255
  # 0 being fully transparant and 255 being fully visable
  # Works with either color and trans a vector of equal length,
  # or one of the two of length 1.
  if (length(color)!=length(trans)&!any(c(length(color),length(trans))==1)){ 
	  stop("Vector lengths not correct")
  }
	if (length(color)==1 & length(trans)>1) color <- rep(color,length(trans)) 
	if (length(trans)==1 & length(color)>1) trans <- rep(trans,length(color))

  num2hex <- function(x) {
	  	hex <- unlist(strsplit("0123456789ABCDEF",split=""))
		return(paste(hex[(x-x%%16)/16+1],hex[x%%16+1],sep="")) 
	}

	rgb <- rbind(col2rgb(color),trans)
	res <- paste("#",apply(apply(rgb,2,num2hex),2,paste,collapse=""),sep="") 
	return(res)
}

```

## SLR Data Generation
In the following examples, I'll generate a dataframe with data based on a simple linear regression model. Generally, the procedure here is to create a function that will populate a dataframe, with each variable following the prescribed formula. Then when we have the data, running lm on it to get the coefficient estimates.

The linear model is as follows and the usual assumptions (A1-4) hold:
\(Y = \beta_0 + \beta_1 X_1 + \epsilon\)

  * the sample size, \(n\) is given, e.g. \(n=100\)
  * the distribution of \(X_1\) is given, e.g. \(X_1 \sim U(-1, 1) \)
  * the distribution of the error terms \(\epsilon\) is given, e.g. \(\epsilon \sim t_1\)
  * for simulations, we're also given values of \(\beta_0\) and \(\beta_1\)
  
### Example 1:  \(\epsilon \sim t_1\)
Let's say X follows a uniform distribution in the interval \([-1, 1]\) and we have \(\beta_0 =5\), \(\beta_1 = -2\). Generate data for a SLR model where the errors follow a t distribution with 1 degree of freedom.

Note the difference between the true regression line (following the given values of \(\beta\)) and the fitted regression line, attained from a sample from the data.

```{r errors ~ t}
errors_t_df <- function(n, beta0, beta1, df){
  x <- runif(n, min=-1, max=1)  # built in r function for the uniform distribution
  e <- rt(n, df=df)  # draw n points from the t dist with df degrees of freedom
  y <- beta0 + beta1 * x + e  # this is our linear model spec
  return(data.frame(x=x, y=y))
}

# Create the data
errors_t_model.data = errors_t_df(n=10000, beta0=5, beta1=-2, df=1)
plot(errors_t_model.data$x, errors_t_model.data$y, col=addTrans('orange', 30), ylim=c(-5,20))
# true regression line
abline(a=5, b=-2, col='red', lwd=3)

# fitted regresion line
errors_t_model.sample = errors_t_df(n=50, beta0=5, beta1=-2, df=1)
points(errors_t_model.sample$x, errors_t_model.sample$y, col=addTrans('blue', 200))
errors_t_model.fit = lm(y ~ x, data=errors_t_model.sample)
abline(errors_t_model.fit, col=addTrans('blue', 200))
```


### Example 2:  \(\epsilon \sim \text{Gamma}\)
Same idea here and same parameters. Except now we consider the case where the error term comes from another distribution:

Let \(\eta \sim \Gamma(k=2, \theta=1)\). \(\mathbb{E}(\eta) = k \cdot \theta = 2\).

But we would like the error term to have zero expectation, so we can set \(\epsilon = \eta - 2\) so that \(\mathbb{E}(\epsilon) = \mathbb{E}(\eta - 2) = 2 -2 = 0\).

```{r gamma errors}
errors_gamma_df <- function(n, beta0, beta1, k, theta){
  x <- runif(n, min=-1, max=1)  
  e <- rgamma(n, shape=k, scale=theta) - k * theta
  y <- beta0 + beta1 * x + e  
  return(data.frame(x=x, y=y))
}

# Create the data
errors_gamma_model.data = errors_gamma_df(n=10000, beta0=5, beta1=-2, k=2, theta=1)
plot(errors_gamma_model.data$x, errors_gamma_model.data$y, col=addTrans('orange', 30))
# true regression line
abline(a=5, b=-2, col='red', lwd=3)

# fitted regresion line
errors_gamma_model.sample = errors_gamma_df(n=10, beta0=5, beta1=-2, k=2, theta=1)
points(errors_gamma_model.sample$x, errors_gamma_model.sample$y, col=addTrans('blue', 200))
errors_gamma_model.fit = lm(y ~ x, data=errors_gamma_model.sample)
abline(errors_gamma_model.fit, col=addTrans('blue', 200))
```

## Other distributions for errors
```{r}
## normal:
e <- rnorm(n=100, mean=0, sd=1)
```

